---
title: "Brazil Municipalities SPATIAL DATA ANALYSIS"
author: "BLESSING NEHOHWA"
date: "`r Sys.Date()`"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---

```{r}
##install.packages('tidyverse')
##install.packages('sf')
##install.packages('scales')
##install.packages('ggplot2')
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd ("C:/Users/beemu/OneDrive/SDA")


library(tidyverse)
library(sf)
library(ggplot2)
library(scales)
library(ggsn)
```



## Question 1 
   Produce a map showing the centroids of each municipality in just the state of Säo  
   Paulo, and add the outer boundary of Säo Paulo state.

1a). Open municipalities and filter for Sp state and print selected data within the  
     console

```{r}
municipalities<- read_sf("C:/Users/beemu/OneDrive/SDA/municipalities/municipalities.shp" )
municipalities_SP<-municipalities%>% filter(UF=="SP")
print(municipalities_SP[1:10,])
```


1b).  Generating municipalities centroids and plotting them

```{r warning=FALSE}
municipalities_centroids<-municipalities_SP%>%st_centroid()
municipalities_centroids %>% ggplot()+geom_sf()
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.


1c).Dissolve SP State polygons to create SP state Boarder

```{r}
SP_boarder<-municipalities_SP%>%st_union()
SP_boarder%>% ggplot()+geom_sf()
```


1d).    Plotting SP centroids and SP boarder together

```{r}
SP_boarder%>%ggplot() +geom_sf() +
 geom_sf(data=municipalities_centroids) +
  theme_classic()
```

## Question 2
What is the mean Human Development Index of Municipalities in each state of Brazil?
HDI is the variable from Municipalities data set.

2a).Reading The table

```{r}
br_states <-municipalities[,]

print(br_states)
```

2b). Calculating the mean HDI by state

```{r}
br_states <-municipalities %>% group_by(UF) %>% summarize(IDHM_mean=mean(IDHM_10))

br_states
```

2c). Plotting the result (Mean)

```{r}
br_states %>%ggplot() +
  geom_sf(aes(fill =IDHM_mean)) +
  theme_classic() +
  scale_fill_continuous(low="red",high="green")

```


## Question 3
Produce a polygon/shapefile mapping the area of the municipality ‘Gaucha do Norte’ that is in the indigenous territory “Parque do Xingu”

3a). Import indigenous park shapefile and plot it.
     The projections are the same 

```{r}
indigenous<-read_sf("C:/Users/beemu/OneDrive/SDA/BC250_Terra_Indigena_A/BC250_Terra_Indigena_A.shp") %>% st_transform(4326)
br_states %>% ggplot() +
  geom_sf() +
  geom_sf(data=indigenous,fill="red")

```


3b). Creating a shapefile for Gaucha do Norte e Xingu.

```{r}
Xingu <-indigenous %>% filter(nome=="Parque do Xingu")%>% st_transform(4326)
Gaucha <- municipalities %>% filter(NOME=="GAUCHA DO NORTE") %>% st_transform (4326)
```


3c). Plot the shapefiles to see the overlap
```{r}
Gaucha%>% ggplot() +
  geom_sf(fill="red") +
  geom_sf(data=Xingu,fill="blue",alpha=0.5)
```


3d).Creating shapefile for intersection and plotting it
    Attribute variables are assumed to be spatially constant throughout all geometries
    changed fill colour to Orange
    
```{r}
intersection <- Gaucha %>% st_intersection(Xingu)
intersection %>% ggplot() + geom_sf(fill="black")

```


3e). Plot all together

```{r}
 Gaucha %>% ggplot() +
  geom_sf(fill="red", alpha=0.5) +
  geom_sf(data=Xingu, fill="blue", alpha=0.5) +
  geom_sf(data = intersection, fill="black")

```

3e.)  The  area of the intersection

```{r}
st_area(intersection)
```


## Question 4 
In the state of Acre (AC), which two social housing (MCMV) projects are closest to each other? Create a 10km buffer around each housing project.

4a). Import Housing Shapefile

```{r}

Housing <- read_sf("C:/Users/beemu/OneDrive/SDA/MCMV_new/MCMV_new.shp")
```


4b). Select housing in AC state and plot the result.

```{r}
Housing_AC <- Housing %>% filter(UF=="AC")
br_states %>% filter(UF=="AC") %>% ggplot() +
  geom_sf(fill="light blue") + geom_sf(data = Housing_AC)

```
 
 4c). Calculating distance between housing points and view results
 
```{r}
 distance <- Housing_AC %>%
  st_transform(29189) %>%
  st_distance() %>%
  as.data.frame()

print(distance[1:5,])
```

4d). Calculating buffer around housing points

```{r}
Housing_AC %>% st_transform(29189) %>%
  st_buffer(20000) %>% 
  ggplot() + geom_sf(fill="dark green")
```

4e). Plotting it all Together

```{r}
br_states %>% filter(UF=="AC") %>% ggplot() +
  geom_sf(fill="light blue") + 
  geom_sf(data = (Housing_AC %>% st_transform(29189) %>% st_buffer(20000)), fill = "dark green") +
  geom_sf(data = Housing_AC, color = "red")
```


##Question 5 
Across Brazil, which municipalities have the lowest and highest number of MCMV housing units (UH) in its territory? Create a map of the distribution of total housing units by municipality.

5a). Spatial join between municipalities and housing

```{r}
mun_Housing_units <-municipalities %>% st_join(Housing)
print(mun_Housing_units[1:5,])
```


5b). Calculating total of Housing units in each municipality

```{r}
mun_Housing_units <- mun_Housing_units %>%
  group_by(COD_MUN,NOME) %>%
  summarise(UH=sum(UH,na.rm = T)) %>%
  ungroup()

```


5c). Select municipality with most Housing units

```{r}
mun_Housing_units %>% arrange(-UH) %>% slice(1) %>% pull(NOME)
```

5d) Plotting the distribution of total Housing units by municipality

```{r}

mun_Housing_units %>% ggplot() +
  geom_sf(aes(fill=UH), col=NA) +
  scale_fill_gradient(low="#ccece6",high="dark blue",trans="log") +
  theme_classic()
```

NB: Transformation introduced infinite values in discrete y-axis


##EXTRA: CREATE A VORONOI POLYGON FOR SP STATE

Extra 1a). Installing and Loading needed libraries

```{r}
#install.packages("dismo")
library(deldir)
library (dismo)
```


Extra 1b).   Creating and viewing Voronoi polygon

```{r}
sp_voronoi <-municipalities_centroids %>%
  as("Spatial") %>%
  voronoi() %>%
  st_as_sf()
sp_voronoi %>% ggplot() +
  geom_sf(fill="white")
```


Extra 1c).   Clipping Voronoi to SP State

```{r}
sp_voronoi_clip <- sp_voronoi %>% st_intersection(st_union(SP_boarder))
sp_voronoi_clip %>% ggplot() + geom_sf()
```
Warning: attribute variables are assumed to be spatially constant throughout all geometries.



## EXtra 2: Count points inside a polygon
            Perform operation and view results

```{r}
mun_housin <- municipalities %>% st_join(Housing) %>%
  group_by(COD_MUN) %>%
  count()

mun_housin %>% ggplot() +
  geom_sf(aes(fill=n),col=NA) +
  scale_fill_gradient(low = "#ccece6", high = "purple") +
  theme_classic()
```



